---
title: "230/232 Lab 3"
Authors: "Shane O'Brian, Rosemary Juarez, Liane Chen"
date: "2024-04-17"
output: html_document
---

```{r setup}
#setting up our knitr options
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```

```{r libraries, results="hide"}
#calling in libraries
library(tidyverse)
library(janitor)
library(kableExtra)
```

```{r}
# read in data: 
clim_data <- read_table("clim.txt") %>% 
  clean_names() #cleaning up the name

# Display the first few rows of the dataframe
kable(head(clim_data), "pipe")
#will look bbetter once we run through html
```


## Lab Instructions

For Assignment 3, our group will:

-   Do a simple informal sensitivity analysis of almond yield profit using at least 2 parameters

-   Create 2 graphs of the results - you can decide what is the most meaningful graph

-   Write a short paragraph (in the Rmarkdown document) to summarize you interpretation of your model results (e.g what do 'take away' from your model and sensitivity analysis)


## Sourcing and Running Models
We combined our almond yield calculation and profit model in a new function script called `profit_model.R`. out output will be a dataframe
```{r}
#-----------------------------------
#     Sourcing
#-----------------------------------
source("almond_yield.R")
source("profit_model.R") # contains both almond yield calculation and profit model


almond_yield_and_profit(clim_data, 2000, 100000, 5)

```

## Informal Sensitivity Analysis

```{r}
yields = parm %>% pmap(compute_almond_yield,clim=clim)
```

